Problems Involving Legendre Polynomials

### 2016-02-12 number_theory/hw/hw4.tex
more work done

\item Suppose that $a$ and $b$ are positive integers.
\begin{enumerate}
\item If $a^2\mid b^2$ prove that $a\mid b$. (Hint: Assuming that $a^2m = b^2$, use unique factorization to
prove that $m$ is a perfect square.)
\begin{quote}
Assume that $a^2\mid b^2$. Then $b^2=a^2m$ for some $m\in\ZZ$. By the Fundamental Theorem of Arithmetic,
we have that
$(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^2 =
(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^2m$ for $j,k\in\ZZ$ and $\alpha,\beta\in\NN$.
\end{quote}

\item If $n$ is a positive integer such that $a^n \mid b^n$ can we conclude that $a \mid b$? Either prove
your answer or provide a counterexample.
\begin{quote}

\end{quote}
\end{enumerate}

@@

\item Two primes $p$ and $q$ with $p<q$ are called twin primes if $p+2=q$. If $p$ and $q$ are twin primes
with $3 < p < q$, prove that $6 \mid p + 1$.
\begin{quote}

\end{quote}
\end{enumerate}
\end{document}

>>>>

\item Suppose that $a$ and $b$ are positive integers.
\begin{enumerate}
\item If $a^2\mid b^2$ prove that $a\mid b$. (Hint: Assuming that $a^2m = b^2$, use unique factorization to
prove that $m$ is a perfect square.)
\begin{quote}
Assume that $a^2\mid b^2$. Then $b^2=a^2m$ for some $m\in\ZZ$. By the Fundamental Theorem of Arithmetic,
we have that
$(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^2 =
(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^2m$ for $j,k\in\ZZ$ and $\alpha,\beta\in\NN$.
Then we can say that
\[
{
(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^2
\over
(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^2
}
= m\in\ZZ.
\]
Therefore, $m$ is a perfect square, so we can take its square root. Let $n=\sqrt{m}$. Then
$\sqrt{a^2m}=\sqrt{b^2}$, or $an=b$, so $a|b$.
\end{quote}

\item If $n$ is a positive integer such that $a^n \mid b^n$ can we conclude that $a \mid b$? Either prove
your answer or provide a counterexample.
\begin{quote}
By the same argument used above, we can say
\[
{b^n\over a^n} =
{
(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^n
\over
(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^n
}
= m\in\ZZ.
\]
Therefore, $m$ is a perfect $n^{\text{th}}$ power, so we can say $j=\sqrt[n]{m}$. Then
$\sqrt[n]{b^n}=\sqrt[n]{a^nm}$ simplifies to $b=aj$. Therefore, $a|b$.
\end{quote}
$\triangle$
\end{enumerate}

@@

\item Two primes $p$ and $q$ with $p<q$ are called twin primes if $p+2=q$. If $p$ and $q$ are twin primes
with $3 < p < q$, prove that $6 \mid p + 1$.
\begin{quote}
We can take any three consecutive numbers starting with a prime $p$ where $p+2=q$ is also prime. We know
that any prime number greater than 3 is even, so it is divisible by two. We also know that if we take
three consecutive numbers, one of them has to be divisible by 3, since it will either have a remainder of
0, 1, or 2. Since $p$ is prime and $p+2$ is prime, then $p+1$ must be divisible by three. Therefore
$p+1$ must be divisible by 6, or $6|p+1$.
\end{quote}
\end{enumerate}
\end{document}

### 2016-02-12 partial_differential_equations/leg.tex
more work done

\item We want to find $a_{l}$ in the series
\[
f(x) = \sum_{l=0}^{\infty}a_{l}p_{l}(x)
\]
where
\[
f(x)=
\begin{cases}
x,\ 0\leq x < 1\\
0,\ -1 < x \leq 1
\end{cases}.
\]
We know that we can express $a_{l}$ as
\[
a_{l} = {%
\int_{-1}^{1}f(x)p_{l}(x)\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
Since $f$ is piecewise defined to be $0$ on $-1 < x \leq 1$, we are only dealing with
\[
a_{l} = {%
\int_{0}^{1}xp_{l}(x)\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
We can use integration by parts to solve the numerator. If we let $u = x$ and
$\dv = p_{l}(x)\dx$, we have that
\[
{%
{x\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\Big|_{0}^{1}%
-\int_{0}^{1}{1\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
Let's consider the first term in the numerator:
${x\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\Big|_{0}^{1}$.
In order to evaluate the differentiated term, we need to express the derivative in a closed form.
If we make the substitution $y=x^{2}$, we have
\[
(y-1)^{l} = (y-1)(y^{l-1}+y^{l-2}+\cdots+1).
\]
By back substitution, we have
\begin{align*}
(x^{2}-1)^{l} &= (x^2-1)(x^{2l-2}+x^{2l-4}+\cdots+1)\\
&= (x^2-1)(\sum_{j=0}^{l}x^{(2l-2j)})\\
&= x^{2l+2}-1
\end{align*}

\end{enumerate}

>>>>

\newpage
\item We want to find $a_{l}$ in the series
\[
f(x) = \sum_{l=0}^{\infty}a_{l}p_{l}(x)
\]
where
\[
f(x)=
\begin{cases}
x,\ 0\leq x < 1\\
0,\ -1 < x \leq 1
\end{cases}.
\]
We know that we can express $a_{l}$ as
\[
a_{l} = {%
\int_{-1}^{1}f(x)p_{l}(x)\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
Since $f$ is piecewise defined to be $0$ on $-1 < x \leq 1$, we are only dealing with
\[
a_{l} = {%
\int_{0}^{1}xp_{l}(x)\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
We will begin by analyzing the numerator. We are solving
\[
\int_{0}^{1}xp_{l}(x)\dx.
\]
Using integration by parts gives us
\begin{align*}
&x{1\over2^ll!}{d^{l-1}\over dx^{l-1}}\left[(x^2-1)^l\right]-
{1\over2^ll!}{d^{l-2}\over dx^{l-2}}\left[(x^2-1)^l\right]\Big|_{x=0}^{x=1}\\
&= xp_{l-1}(x)-p_{l-2}(x)\Big|_{x=0}^{x=1}\\
&= \left[1\cdot p_{l-1}(1) - p_{l-2}(1)\right] -\left[0-p_{l-2}(0)\right]\\
&= p_{l-2}(0).
\end{align*}
We can express $p_{l}(0)$ as
\[
p_{l}(0) =
\begin{cases}
{(-1)^{l/2}\over2^l}{l\choose l/2}\quad x\in\mathbb{E}\\
0\qquad\qquad\quad\ x\in\mathbb{O}
\end{cases}.
\]
We will consider even $l$ here, since all odd $l$ will produce a 0. We can now express $a_{l}$ as
\begin{align*}
a_{l} &= {(2l+1)p_{l-2}(0)\over2}\\
&= {4(2l+1)\over l!}.
\end{align*}
The Fourier series is now
\[
\sum{4(2l+1)\over l!}p_{l}(x)
\]
for even $l$.
\end{enumerate}
